---
title: "02_clonal_diversity_analysis"
author: "Enric Vercher"
date: "2023-11-07"
output: html_document
---

```{r}
# load libraries
suppressPackageStartupMessages(library(airr))
suppressPackageStartupMessages(library(alakazam))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(dowser))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(scoper))
suppressPackageStartupMessages(library(shazam))
suppressPackageStartupMessages(library(tigger))
```

### Setting the clonal distance threshold with SHazaM


```{r}
m.44.s <- airr::read_rearrangement('C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/m.44_genotyped.tsv')
m.45.s <- airr::read_rearrangement('C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/m.45_genotyped.tsv')
m.48.s <- airr::read_rearrangement('C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/m.48_genotyped.tsv')
m.49.s <- airr::read_rearrangement('C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/m.49_genotyped.tsv')
m.50.s <- airr::read_rearrangement('C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/m.50_genotyped.tsv')
m.52.s <- airr::read_rearrangement('C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/m.52_genotyped.tsv')
```


## Calculate distance 

```{r}
m.44.s <- distToNearest(m.44.s, model = "ham", normalize = "len",
                    vCallColumn = "v_call_genotyped", nproc = 4)
m.45.s <- distToNearest(m.45.s, model = "ham", normalize = "len",
                    vCallColumn = "v_call_genotyped", nproc = 4)
m.48.s <- distToNearest(m.48.s, model = "ham", normalize = "len",
                    vCallColumn = "v_call_genotyped", nproc = 4)
m.49.s <- distToNearest(m.49.s, model = "ham", normalize = "len",
                    vCallColumn = "v_call_genotyped", nproc = 4)
m.50.s <- distToNearest(m.50.s, model = "ham", normalize = "len",
                    vCallColumn = "v_call_genotyped", nproc = 4)
m.52.s <- distToNearest(m.52.s, model = "ham", normalize = "len",
                    vCallColumn = "v_call_genotyped", nproc = 4)
```


# Files with distance for the slurm cluster

```{r}
write.table(m.44.s, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/mouse_44.genotyped.dist.tsv", sep = "\t", row.names = FALSE)
write.table(m.45.s, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/mouse_45.genotyped.dist.tsv", sep = "\t", row.names = FALSE)
write.table(m.48.s, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/mouse_48.genotyped.dist.tsv", sep = "\t", row.names = FALSE)
write.table(m.49.s, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/mouse_49.genotyped.dist.tsv", sep = "\t", row.names = FALSE)
write.table(m.50.s, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/mouse_50.genotyped.dist.tsv", sep = "\t", row.names = FALSE)
write.table(m.52.s, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/mouse_52.genotyped.dist.tsv", sep = "\t", row.names = FALSE)
```


## Calculate threshold con FindThreshold

```{r}

threshold_output_m44 <- shazam::findThreshold(m.44.s$dist_nearest,
                                  method = "gmm", model = "gamma-norm",
                                  cutoff = "user", spc = 0.995)
threshold_m44 <- threshold_output_m44@threshold
threshold_m44
```

```{r}
plot(threshold_output_m44, binwidth = 0.02, silent = T) +
  theme(axis.title = element_text(size = 18))
```


```{r}
threshold_output_m45 <- shazam::findThreshold(m.45.s$dist_nearest,
                                  method = "gmm", model = "gamma-norm",
                                  cutoff = "user", spc = 0.995)
threshold45 <- threshold_output_m45@threshold
threshold45
```

```{r}
plot(threshold_output_m45, binwidth = 0.02, silent = T) +
  theme(axis.title = element_text(size = 18))
```

```{r}
threshold_output_m48 <- shazam::findThreshold(m.48.s$dist_nearest,
                                  method = "gmm", model = "gamma-norm",
                                  cutoff = "user", spc = 0.995)
threshold48 <- threshold_output_m48@threshold
threshold48

plot(threshold_output_m45, binwidth = 0.02, silent = T) +
  theme(axis.title = element_text(size = 18))
```


```{r}
threshold_output_m49 <- shazam::findThreshold(m.49.s$dist_nearest,
                                  method = "gmm", model = "gamma-norm",
                                  cutoff = "user", spc = 0.995)
threshold49 <- threshold_output_m49@threshold
threshold49

plot(threshold_output_m49, binwidth = 0.02, silent = T) +
  theme(axis.title = element_text(size = 18))
```

```{r}
threshold_output_m50 <- shazam::findThreshold(m.50.s$dist_nearest,
                                  method = "gmm", model = "gamma-norm",
                                  cutoff = "user", spc = 0.995)
threshold50 <- threshold_output_m50@threshold
threshold50

plot(threshold_output_m50, binwidth = 0.02, silent = T) +
  theme(axis.title = element_text(size = 18))
```


```{r}
threshold_output_m52 <- shazam::findThreshold(m.52.s$dist_nearest,
                                  method = "gmm", model = "gamma-norm",
                                  cutoff = "user", spc = 0.995)
threshold52 <- threshold_output_m52@threshold
threshold52

plot(threshold_output_m52, binwidth = 0.02, silent = T) +
  theme(axis.title = element_text(size = 18))
```

## Spectral-based clonal clustering (SCOPer)

Como no son bimodales, identificamos los clones por spectral clustering

### mouse_44

```{r}
# Clonal assignment using identical nucleotide sequences
m.44.scoper <- identicalClones(m.44.s, method="nt")
m.44.scoper_db <- as.data.frame(m.44.scoper)
```

```{r}
# Plot a histogram of inter clonal distances
plot(m.44.scoper, binwidth=0.02)
```

```{r}
# Clonal assignment using hierarchical clustering
results_m44 <- hierarchicalClones(m.44.s, threshold=0.15)
```

```{r}
# Plot a histogram of inter and intra clonal distances
plot(results_m44, binwidth=0.02)
```

```{r}
# Clonal assignment using the spectral clustering method novj
results_m44b <- spectralClones(m.44.s, method="novj")
# Plot a histogram of inter and intra clonal distances
plot(results_m44b, binwidth=0.02)
```

### mouse_45

```{r}
# Clonal assignment using identical nucleotide sequences
m.45.scoper <- identicalClones(m.45.s, method="nt")
m.45.scoper_db <- as.data.frame(m.45.scoper)
```

```{r}
# Plot a histogram of inter clonal distances
plot(m.45.scoper, binwidth=0.02)
```

```{r}
# Clonal assignment using hierarchical clustering
results_m45 <- hierarchicalClones(m.45.s, threshold=0.15)
```

```{r}
# Plot a histogram of inter and intra clonal distances
plot(results_m45, binwidth=0.02)
```

Este no lo usamos porque findhthrehold si que ha funcionado en este

```{r}
# Clonal assignment using the spectral clustering method novj
results_m45b <- spectralClones(m.45.s, method="novj")
# Plot a histogram of inter and intra clonal distances
plot(results_m45b, binwidth=0.02)
```

### Mouse_48

```{r}
# Clonal assignment using identical nucleotide sequences
m.48.scoper <- identicalClones(m.48.s, method="nt")
m.48.scoper_db <- as.data.frame(m.48.scoper)
```

```{r}
# Plot a histogram of inter clonal distances
plot(m.48.scoper, binwidth=0.02)
```

```{r}
# Clonal assignment using hierarchical clustering
results_m48 <- hierarchicalClones(m.48.s, threshold=0.15)
```

```{r}
# Plot a histogram of inter and intra clonal distances
plot(results_m48, binwidth=0.02)
```

```{r}
# Clonal assignment using the spectral clustering method novj
results_m48b <- spectralClones(m.48.s, method="novj")
# Plot a histogram of inter and intra clonal distances
plot(results_m48b, binwidth=0.02)
```

### Mouse_49

```{r}
# Clonal assignment using identical nucleotide sequences
m.49.scoper <- identicalClones(m.49.s, method="nt")
m.49.scoper_db <- as.data.frame(m.49.scoper)
```

```{r}
# Plot a histogram of inter clonal distances
plot(m.49.scoper, binwidth=0.02)
```

```{r}
# Clonal assignment using hierarchical clustering
results_m49 <- hierarchicalClones(m.49.s, threshold=0.15)
```

```{r}
# Plot a histogram of inter and intra clonal distances
plot(results_m49, binwidth=0.02)
```

```{r}
# Clonal assignment using the spectral clustering method novj
results_m49b <- spectralClones(m.49.s, method="novj")
# Plot a histogram of inter and intra clonal distances
plot(results_m49b, binwidth=0.02)
```

### Mouse_50

```{r}
# Clonal assignment using identical nucleotide sequences
m.50.scoper <- identicalClones(m.50.s, method="nt")
m.50.scoper_db <- as.data.frame(m.50.scoper)
```

```{r}
# Plot a histogram of inter clonal distances
plot(m.50.scoper, binwidth=0.02)
```

```{r}
# Clonal assignment using hierarchical clustering
results_m50 <- hierarchicalClones(m.50.s, threshold=0.15)
```

```{r}
# Plot a histogram of inter and intra clonal distances
plot(results_m50, binwidth=0.02)
```

```{r}
# Clonal assignment using the spectral clustering method novj
results_m50b <- spectralClones(m.50.s, method="novj")
# Plot a histogram of inter and intra clonal distances
plot(results_m50b, binwidth=0.02)
```

### Mouse_52

```{r}
# Clonal assignment using identical nucleotide sequences
m.52.scoper <- identicalClones(m.52.s, method="nt")
m.52.scoper_db <- as.data.frame(m.52.scoper)
```

```{r}
# Plot a histogram of inter clonal distances
plot(m.52.scoper, binwidth=0.02)
```

```{r}
# Clonal assignment using hierarchical clustering
results_m52 <- hierarchicalClones(m.52.s, threshold=0.15)
```

```{r}
# Plot a histogram of inter and intra clonal distances
plot(results_m52, binwidth=0.02)
```

```{r}
# Clonal assignment using the spectral clustering method novj
results_m52b <- spectralClones(m.52.s, method="novj")
# Plot a histogram of inter and intra clonal distances
plot(results_m52b, binwidth=0.02)
```

```{r}
# individual threshold con findThreshold
thr44 <- 0.0697
thr45 <- 0.0353
thr48 <- 0.0274
thr49 <- 0.1111
thr50 <- 0.0633
thr52 <- 0.1224
mean_find <- mean(c(0.0697,0.0353,0.0274,0.1111,0.0633,0.1224))
median_find <- median(c(0.0697,0.0353,0.0274,0.1111,0.0633,0.1224))
mean_find
median_find
```

No me fio de estos resultados, no parecen dividir las montañas. Sin embargo, el findthreshold ha sido capaz de entoncontrar un umbral al que cortar la jerarquía. 

```{r}
# individual threshold con spectral clones
thr44 <- 0.18
thr45 <- 0.19 #del findhthrehold
thr48 <- 0.15 #deducido, no lo encuentra
thr49 <- 0.31
thr50 <- 0.3
thr52 <- 0.3
mean_spectral <- mean(c(0.18,0.19,0.15,0.31,0.3,0.3))
median_spectral <- median(c(0.18,0.19,0.15,0.31,0.3,0.3))
mean_spectral
median_spectral
```

```{r}
results_m44b <- as.data.frame(results_m44b)
results_m45b <- as.data.frame(results_m45b)
results_m48b <- as.data.frame(results_m48b)
results_m49b <- as.data.frame(results_m49b)
results_m50b <- as.data.frame(results_m50b)
results_m52b <- as.data.frame(results_m52b)
```

Sin embargo, con el spectral clones, no hemos sido capaces de encontrar un threshold para el mouse_48 (pocas células).

## Altogether

The nearest-neighbor distance distribution is not always bimodal. In this case, if the data have multiple subjects, we can calculate the nearest neighbor distances across subjects to initialize the Gaussian fit parameters of the nearest-neighbor distance of inter (between) clones distribution. 

The nearest neighbor distances across subjects can be calculated by specifying the parameter `cross` in the function `distToNearest`. And then when we call function `findThreshold`, Gaussian fit parameters can be initialized by setting parameter `cross = dist_crossSubj$cross_dist_nearest`.

In the above data, there are 6 subjects. We will want to make make sure that the cross subject `distToNearest` values are valid. To calculate this do the following:

```{r}
m.44.s$subject <- "mouse_44"
m.45.s$subject <- "mouse_45"
m.48.s$subject <- "mouse_48"
m.49.s$subject <- "mouse_49"
m.50.s$subject <- "mouse_50"
m.52.s$subject <- "mouse_52"
```

```{r}
all_mice_analysis <- rbind(m.44.s, m.45.s, m.48.s, m.49.s, m.50.s, m.52.s)
```

```{r}
#son lo mismo
#Use genotyped V assignments, Hamming distance, and normalize by junction length
dist_nearest <- distToNearest(filter(all_mice_analysis, locus == "IGH"), nproc = 1,model="ham", normalize ="len",vCallColumn = "v_call_genotyped") #lo mismo que distToNearest(all_mice_analysis)
dist_nearest <- distToNearest(all_mice_analysis,nproc = 1,model="ham", normalize ="len",vCallColumn = "v_call_genotyped")
# calculate cross subjects distribution of distance to nearest
dist_crossSubj <- distToNearest(filter(all_mice_analysis, locus == "IGH"), 
                                nproc = 1, cross = "subject")

# find threshold for cloning automatically and initialize the Gaussian fit parameters of the nearest-neighbor 
# distance of inter (between) clones using cross subjects distribution of distance to nearest
threshold_output <- findThreshold(dist_nearest$dist_nearest, 
                                  method = "gmm", model = "gamma-norm", 
                                  cross = dist_crossSubj$cross_dist_nearest,
                                  cutoff = "user", spc = 0.995)
threshold <- threshold_output@threshold
threshold
```
Este valor (0.05362) es similar a la mediana del conjunto de los thresholds por ratón indivual.

# Plot de todos juntos

```{r}
plot(threshold_output, binwidth = 0.02, 
     cross = dist_crossSubj$cross_dist_nearest, silent = T) + 
  theme(axis.title = element_text(size = 18))
```

#Si los hago por separado y saco la media y mediana

```{r}
# Lista para almacenar los umbrales
thresholds <- c()

# Iteración para cada ratón
for (subject in c("mouse_45", "mouse_48", "mouse_49", "mouse_50", "mouse_52")) {
    # Cálculo del umbral para cada ratón
    dist_nearest <- distToNearest(dplyr::filter(all_mice_analysis, locus == "IGH", subject == subject))
    threshold_output <- findThreshold(dist_nearest$dist_nearest, method = "gmm", model = "gamma-norm", cutoff = "user", spc = 0.995)
    threshold <- threshold_output@threshold
    
    # Almacenar el umbral en la lista
    thresholds <- c(thresholds, threshold)
}

# Calcular la media y la mediana de los umbrales
mean_threshold <- mean(thresholds)
median_threshold <- median(thresholds)

# Mostrar los resultados
mean_threshold
median_threshold
```


# Si hacemos la asignación clonal con Scoper

```{r}
# call clones using hierarchicalClones
results <- hierarchicalClones(all_mice_analysis,
                              cell_id = "cell_id",
                              threshold = 0.05357612,
                              method="nt",
                              only_heavy = TRUE, split_light = FALSE,
                              summarize_clones = FALSE,
                              fields = "subject")
```

Como hemos eliminado las cadenas ligeras, no las encuentra.


# Exportar los datos.

```{r}
all_mice_genotyped <- airr::read_rearrangement('C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/02_Genotipado/all_mice_genotyped.tsv')
```

Esto no lo hemos hecho. La asignación clonal la hacemos con change-o en el cluster con los scripts DefineClones.py y CreateGermlines.py


```{r}
thr_spectr <- spectralClones(all_mice_genotyped, method="novj")
# Plot a histogram of inter and intra clonal distances
plot(thr_spectr, binwidth=0.02)
```

```{r}
# call clones using hierarchicalClones 
All_mice_clustering <- hierarchicalClones(all_mice_genotyped, cell_id = 'cell_id', 
                              threshold = 0.28, only_heavy = TRUE, 
                              split_light = FALSE, summarize_clones = FALSE)
```

```{r}
# calculate and plot the rank-abundance curve
plot(estimateAbundance(All_mice_clustering), colors = "steelblue", silent = T) + 
  theme(axis.title = element_text(size = 18))
```


```{r}
# call clones using hierarchicalClones 
All_mice_clustering2 <- hierarchicalClones(all_mice_analysis, cell_id = 'cell_id', 
                              threshold = 0.053, only_heavy = TRUE, 
                              split_light = FALSE, summarize_clones = FALSE)
```

```{r}
# calculate and plot the rank-abundance curve
plot(estimateAbundance(All_mice_clustering2), colors = "steelblue", silent = T) + 
  theme(axis.title = element_text(size = 18))
```


```{r}
write.table(All_mice_clustering, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/03_Clonal_diversity/All_mice_clusteringt.tsv", sep = "\t", row.names = FALSE)
```


```{r}
write.table(results_m44b, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/03_Clonal_diversity/m44_clonal_assgiment.tsv", sep = "\t", row.names = FALSE)
write.table(results_m45b, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/03_Clonal_diversity/m45_clonal_assgiment.tsv", sep = "\t", row.names = FALSE)
write.table(results_m48b, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/03_Clonal_diversity/m48_clonal_assgiment.tsv", sep = "\t", row.names = FALSE)
write.table(results_m49b, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/03_Clonal_diversity/m49_clonal_assgiment.tsv", sep = "\t", row.names = FALSE)
write.table(results_m50b, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/03_Clonal_diversity/m50_clonal_assgiment.tsv", sep = "\t", row.names = FALSE)
write.table(results_m52b, "C:/Users/d940401/Desktop/Analisis_immcantation/Analisis_Enric/Datos_Enric/Datos_TFM/00_Datos_filtrados_slurm/03_Clonal_diversity/m52_clonal_assgiment.tsv", sep = "\t", row.names = FALSE)
```

El resto lo haremos en el cluster:

```{r}
# Create a data frame with the mouse numbers and threshold values
#Esto para el clustering con ratones individuales
predicted_thresholds <- data.frame(
  mouse_nums = c("mouse_44", "mouse_45", "mouse_48", "mouse_49", "mouse_50", "mouse_52"),
  thresholds = c(0.0697, 0.0353, 0.0274, 0.1111, 0.0633, 0.1224)
)

# Write the data frame to a CSV file
write.csv(predicted_thresholds, "predicted_thresholds.csv", row.names = FALSE)
```

Lo haremos todo junto, ya que luego la asignación clonal no queda bien ( por ejemplo dos clones "1" con diferentes secuencias CDR3).













